---
title: "Simulation workflow examples"
author: "Carles Milà"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: pdf_document
---

```{r rmd options, echo=FALSE}
knitr::opts_chunk$set(fig.align = 'center', message = FALSE, warning = FALSE,
                      echo=FALSE)
```


```{r libraries}
library("sf")
library("raster")
library("rasterVis")
library("gstat")
library("cowplot")
library("tidyverse")
library("doParallel")

set.seed(1234)
source("code/sim1_functions.R")
```

This document contains the code to run the examples and figures of sections 2.2 and 2.3 of the manuscript, where the two simulation studies are presented; as well as supporting figure S1, where the variograms used for the covariates field simulation (simulation 1) and an example realisation of each of them are shown.


# Simulation 1: Random Fields

## 1. Grid creation

```{r grid creation, fig.width=8, fig.height=4}
# Create an empty results object
res <- data.frame()

# Create grids (raster and point format) and sampling area
rast_grid <- raster(ncols=300, nrows=100, xmn=0, xmx=300, ymn=0, ymx=100)
point_grid <- st_as_sf(rasterToPoints(rast_grid, spatial = TRUE))
inter_area <- matrix(c(0,0,100,0,100,100,0,100,0,0), ncol=2, byrow=TRUE)
inter_area <- st_sfc(st_polygon(list(inter_area)))
inter_grid <- point_grid[st_intersects(point_grid, inter_area, sparse=FALSE),]
extra_area <- matrix(c(200,0,300,0,300,100,200,100,200,0), ncol=2, byrow=TRUE)
extra_area <- st_sfc(st_polygon(list(extra_area)))
extra_grid <- point_grid[st_intersects(point_grid, extra_area, sparse=FALSE),]

# Plot
both_areas <- bind_rows(mutate(st_sf(geom=inter_area),
                               Area="Sampling and interpolation"),
                        mutate(st_sf(geom=extra_area), Area="Extrapolation")) %>%
  mutate(Area=fct_inorder(as.factor(Area)))
ggplot() +
  geom_sf(data=point_grid, pch=".") +
  geom_sf(data=both_areas, aes(fill=Area, colour=Area), alpha=0.3) +
  scale_fill_viridis_d(end=0.8) +
  scale_colour_viridis_d(end=0.8) +
  theme_bw() +
  theme(legend.position = "bottom")
# ggsave("figures/sim1_areas.png", width=6, height=3, dpi=500)
```

## 2. Covariate fields simulation

```{r covariate simulation, fig.width=8, fig.height=5}
# Define semi-variogram
cov_mod <- vgm(model="Sph", psill=1, range=20, nugget=0)
cov_mod <- gstat(formula=z~1, dummy=TRUE, beta=0, model=cov_mod, nmax=100)

# Simulate 20 covariates and a noise field and stack
cov_points <- quiet(predict(cov_mod, point_grid, nsim=21))
cov_stack <- rasterise_stack(cov_points, 1:20, paste0("cov", 1:20))
noise_stack <- rasterise_stack(cov_points, 21, "snoise")

# png("figures/sim1_covfields.png", width=2000, height=1500, res=300)
rasterVis::levelplot(cov_stack, layout=c(4, 5), scales=list(draw=FALSE))
# dev.off()
```

## 3. Outcome creation

```{r outcome creation, fig.width=8, fig.height=3}
# Continuous outcome using van der Laan's formula
out <- cov_stack$cov1*cov_stack$cov2 + cov_stack$cov10^2 -
  cov_stack$cov3*cov_stack$cov17 - cov_stack$cov15*cov_stack$cov4 +
  cov_stack$cov9*cov_stack$cov5 + cov_stack$cov19 -
  cov_stack$cov20^2 + cov_stack$cov9 * cov_stack$cov8
raw_outcome <- out

# Prepare random noise
rnoise <- raster(ncols=300, nrows=100, xmn=0, xmx=300, ymn=0, ymx=100)
vals <- rnorm(100*300, sd=1)
rnoise <- setValues(rnoise, vals)

# Add noise variables
out <- out + rnoise + noise_stack
names(out)  <- "outcome"

# Plot
plot_rast <- stack(raw_outcome, rnoise, noise_stack, out)
plot_names <- c("Y (no noise)", "Random noise", "Spatial noise", "Y (with noise)")
# png("figures/sim1_outcome.png", width=1500, height=750, res=300)
rasterVis::levelplot(plot_rast, layout=c(2, 2), scales=list(draw=FALSE),
                     names.attr=plot_names)
# dev.off()
```

## 4. Sample simulation

```{r sample simulation}
# Create a grid of sampling possibilities
psamples <- tibble(expand.grid(n=c(100,200,300),
                               distribution=c("sregular","wregular","random",
                                              "wclust", "sclust")))
psamples$geometry <- map2(psamples$n, psamples$distribution,
                          sim1_samples, sarea=inter_area)
psamples <- unnest(psamples, cols = c(geometry)) 
psamples$distribution <- gsub("sreg", "strong reg", psamples$distribution)
psamples$distribution <- gsub("sclust", "strong clust", psamples$distribution)
psamples$distribution <- gsub("clust", "clustered", psamples$distribution)
psamples$distribution <- gsub("w", "weak ", psamples$distribution)
psamples$distribution <- fct_inorder(as.factor(psamples$distribution))

# Plot
ggplot() +
  geom_sf(data=inter_area, alpha=0.1) +
  geom_sf(data=psamples, aes(geometry = geometry), size=0.1) +
  facet_grid(n~distribution) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
# ggsave("figures/sim1_samples.png", dpi=300,ºº height = 4, width=7)
```

## 5. Model fit and evaluation

```{r model fit}
set.seed(1234)
examplesim <- sim_fields(range=20, n_train=200, sample_dist="random")
examplesim <- as.data.frame(examplesim[4:length(examplesim)]) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  mutate(statistic = strsplit(name, "_")[[1]][1],
         method = strsplit(name, "_")[[1]][2]) %>%
  ungroup() %>%
  dplyr::select(-name) %>%
  pivot_wider(names_from = "statistic", values_from = "value")
knitr::kable(examplesim, digits=2)
rm(list=ls())
```

\newpage

# Simulation 2: Virtual species

```{r species}
# Sim2 functions
source("code/sim2_functions.R")

# Read data
spoly <- quiet(st_read(dsn="data/species_vdata.gpkg", layer="sampling_polygon"))
wclim <- stack("data/species_stack.grd")

# Simulate points
set.seed(1234)
psamples1 <- mutate(sim2_samples(100, "sregular", spoly), distr="Strong regular")
psamples2 <- mutate(sim2_samples(100, "wregular", spoly), distr="Weak regular")
psamples3 <- mutate(sim2_samples(100, "random", spoly), distr="Random")
psamples4 <- mutate(sim2_samples(100, "wclust", spoly), distr="Weak clustered")
psamples5 <- mutate(sim2_samples(100, "sclust", spoly), distr="Strong clustered")
psamples <- bind_rows(psamples1, psamples2, psamples3, psamples4, psamples5)
psamples$distr <- fct_inorder(psamples$distr)

# Plot landscape
p1 <- rasterVis::levelplot(stretch(wclim, minv=0, maxv=1), 
                           layout=c(4, 5), margin=F, scales=list(draw=FALSE))

# Plot samples
p2 <- ggplot() +
  geom_sf(data=spoly, alpha=0.2, colour="grey50") +
  geom_sf(data=psamples, size=0.2) +
  facet_wrap(~distr, ncol=2) + 
  theme_bw()

p12 <- plot_grid(p1, p2, nrow=1, labels=c("A", "B"), rel_widths = c(1.2, 1))
p12
# save_plot("figures/sim2_workflow.png", p12, base_height=5, base_asp=1.7)
rm(list=ls())
```

# Appendix: variograms and example realisations of simulation 1

```{r variog examples, fig.width=9, fig.height=8}
source("code/sim1_functions.R")
variog_byrange <- function(range){
  
  # Define and plot variogram
  vgm_mod <- vgm(model="Sph", psill=1, range=range, nugget=0)
  gstat_mod <- gstat(formula=z~1, dummy=TRUE, beta=0, model=vgm_mod, nmax=100)
  
  # Create grids (raster and point format) and sampling area
  rast_grid <- raster(ncols=300, nrows=100, xmn=0, xmx=300, ymn=0, ymx=100)
  point_grid <- st_as_sf(rasterToPoints(rast_grid, spatial = TRUE))
  cov_points <- quiet(predict(gstat_mod, point_grid, nsim=1))
  cov_rast <- rasterise_stack(cov_points, 1, "cov")
  
  # Return objects
  list(vgm_mod = vgm_mod, cov_rast = cov_rast)
}

# Create variograms and examples
set.seed(1234)
variog_examples <- map(c(1,10,20,30,40), variog_byrange)

# Plot variograms
p1 <- plot_grid(plot(variog_examples[[1]]$vgm_mod, cutoff=60),
                plot(variog_examples[[2]]$vgm_mod, cutoff=60),
                plot(variog_examples[[3]]$vgm_mod, cutoff=60),
                plot(variog_examples[[4]]$vgm_mod, cutoff=60),
                plot(variog_examples[[5]]$vgm_mod, cutoff=60),
                nrow=5, ncol=1)

# Plot random fields
stack_rf <- stack(variog_examples[[1]]$cov_rast, variog_examples[[2]]$cov_rast,
                  variog_examples[[3]]$cov_rast, variog_examples[[4]]$cov_rast,
                  variog_examples[[5]]$cov_rast)
p2 <- rasterVis::levelplot(stack_rf, layout=c(1, 5), names.attr=rep("", 5),
                           scales=list(draw=FALSE), margin=T,
                           par.strip.text=list(cex=0), between = list(x=0, y=1.1))
p2$par.settings$layout.heights[c("top.padding") ] <- 0.5

# Create final figure
p3 <- plot_grid(p1, p2, ncol=2, rel_widths = c(1, 1.3))
p3
# save_plot("figures/example_variograms.png", p3, base_height = 8, base_asp = 1.1)
```

